### 2018-08-27 SEDs/notes.tex
starting new repo

\documentclass[11pt,a4paper]{article}
\input{format}

\begin{document}

   \title{Very high-$z$ Quasars: SED notes}
 \date{\today}
\maketitle

\section{Notes from \citet{Roseboom2013}}
There is clear evidence that all powerful AGN are surrounded by a
geometrically thick distribution of optically thick material 
\citep{Rowan-Robinson 1977, Lawrence_Elvis1982, Antonucci_Miller1985, 
Edelson1987, Elvis1994, Richards2006b}.

\section{Caballero et al. 2016}
%% Caballero_2016_MNRAS_463_2064
\citet{Caballero2016}

\begin{tabular}{ l l l l }
\hline \hline
measure	   &	unit	&	typical ranges		&	e.g. figure in R06b  \\
\hline
f$_{\nu}$		    &    uJy		  & from 0.1 to $\sim$100,000d &     Fig 9 \\
log(f$_{\nu}$)	    & ergs/s/cm2/Hz  &	-32 to -24  	     	&	Fig 10, top panel \\
$\nu$f$_{\nu}$     &	ergs/s/cm2        & 	-16 to -12		&	Fig 11, bottom panel  \\
$\nu$L$_{\nu}$     &  ergs/s	           &  43 to  47		&	Fig 12. \\
\hline \hline
\end{tabular}

\bibliographystyle{mn2e}
\bibliography{/cos_pc19a_npr/LaTeX/tester_mnras}

\end{document}