### 2018-08-27 tex/VHzQ_colorcolor/colorplot.tex
README.md and tex directory added

\documentclass[11pt,a4paper]{article}
\input{format}

\begin{document}

   \title{Colour-colour plots for VH$z$Q Selection}
 \date{\today}
\maketitle

\section{(z-J) vs. (J-W1)}
    \begin{figure}
      \centering
      % trim={<left> <lower> <right> <upper>}
      \includegraphics[width=16.0cm, trim={0.5cm 0 0 0},clip]
      {Wang2017_Fig1.jpeg}
      \caption[]{\citet{Wang2017}}
      \label{fig:Wang2017}
    \end{figure}

\section{(z-y) vs. (y-J)}

    \begin{figure}
      \begin{center}
      \includegraphics[width=16.0cm,  trim={1.5cm 0 0 0},clip]
      {Mazzucchelli_2017_Fig1.jpeg}
     \caption[]{\citet{Mazzucchelli2017}}
      \label{fig:Mazzucchelli2017}
      \end{center}
    \end{figure}

\section{(i-z) vs. (z-y)}
    \begin{figure}
      \centering
      \includegraphics[width=16.0cm,  trim={1.5cm 0 0 0},clip]
      {Matsuoka_2016_Fig1.jpeg}
      \caption[]{\citet{Matsuoka2016}}
      \label{fig:Matsuoka2016}
    \end{figure}

    \begin{figure}
      \centering
      \includegraphics[width=16.0cm,  trim={1.5cm 0 0 0},clip]
      {Matsuoka_2018_Fig11.jpeg}
      \caption[]{\citet{Matsuoka2018}}
      \label{fig:Matsuoka2018}
    \end{figure}

\section{(z-y) vs. (i-z)}

    \begin{figure}
      \centering
%      \includegraphics[width=16.0cm,  trim={1.5cm 0 0 0},clip, angle=90]
      \includegraphics[width=16.0cm,  trim={0.0cm 0 0 0},clip, angle=0]
      {Banados_2014_Fig1.jpeg}
      \caption[]{\citet{Banados2014}. 
        Right: color–color diagram showing the criteria used to select quasar
        candidates (long-dashed line, upper left corner). The dot-dashed line
        is used for candidates with upper limits in the iP1 band. The thick
        black line shows the expected color of the quasar template from
        \citet{Decarli2010} redshifted from $z=5.0$ to $z=6.5$ in steps of
        $\Delta z = 0.1$ (see left panel). The M/L/T dwarfs from \citet{Dupuy_Liu2012}  that have a PS1 counterpart are shown with stars. Blue squares are
        published quasars at 5.7 < z < 6.4 satisfying our S/N and coverage
        criteria. Fourteen of them were included in our candidate list. Green
        triangles represent the PS1 colors of unpublished spectroscopically
        confirmed z ∼ 6 quasars (S. J. Warren et al., in preparation). Three
        of them were part of our candidate list (see Tables 4 and 5). The red
        diamonds are the new quasars presented in this paper (see Table
        2). They are labeled with the following letters: a = PSO
        J340.2041–18.6621 (z = 6.00), b = PSO J007.0273+04.9571 (z = 5.99), c
        = PSO J037.9706–28.8389 (z = 5.99), d = PSO J187.3050+04.3243 (z =
        5.89), e = PSO J213.3629–22.5617 (z = 5.88), f = PSO J183.2991–12.7676
        (z = 5.86), g = PSO J210.8722–12.0094 (z = 5.84), h = PSO
        J215.1514–16.0417 (z = 5.73), and i = PSO J045.1840–22.5408 (z =
        5.70).
}
      \label{fig:Banados2014}
    \end{figure}

    \begin{figure}
      \centering
%      \includegraphics[width=16.0cm,  trim={1.5cm 0 0 0},clip, angle=90]
      \includegraphics[width=16.0cm,  trim={0.0cm 0 0 0},clip, angle=0]{f1_ps1_color_selection.pdf}
      \caption[]{\citet{Banados2016}.} 
      \label{fig:Banados2016}
    \end{figure}

\section{(W1 - W2) vs. (z - W2)}
    \begin{figure}
      \centering
      \includegraphics[width=16.0cm,  trim={0.0cm 0 0 0},clip, angle=0]{Carnall_2015_Fig1.jpeg}
      \caption[]{\citet{Banados2016}.} 
      \label{fig:Banados2016}
    \end{figure}

\newpage
    \begin{figure}
      \centering
      \includegraphics[width=16.0cm,  trim={0.0cm 0 0 0},clip, angle=0]{f4_yj_selection.pdf}
      \includegraphics[width=16.0cm,  trim={0.0cm 0 0 0},clip, angle=0]{f6_ntt_colors.pdf}
      \includegraphics[width=16.0cm,  trim={0.0cm 0 0 0},clip, angle=0]{f5_wise_prioritization.pdf}
      \caption[]{\citet{Banados2016}.} 
      \label{fig:Banados2016_3panels}
    \end{figure}

\bibliographystyle{mn2e}
\bibliography{/cos_pc19a_npr/LaTeX/tester_mnras}

\end{document}