### 2019-08-12 tex/MNRAS_submission/rv1/referees_reply.txt
rv1/referees_reply.txt updated

p. 1: You may want to end the first sentence of the right column of the Intro
with a statement that at z>5, relativistic time dilation also stretches the
photometric variability of the accretion disk as observed.

To Do by NPR

p. 2, Fig. 1: Is it useful here to plot both filter sets if they are sufficiently
different? Please provide larger tick-marks for better readability in any case.

Hmmm. I’m not sure what to do here; it’s going to ruin the clarity of the figure
having both ZYJHK filter sets, but, they are different.
I think adding an explicit link to:
       d80b2t/VHzQ/SEDs/filters_vs_QSOstars_UKIRT.png
and
       d80b2t/VHzQ/SEDs/filters_vs_QSOstars_VISTA.png
is the way forward here.

p. 3, Fig. 3, caption or text: You should briefly comment on why the quasar
distribution has this gap in N(z) at z~5.5?. Is this because of the relativity
sensitivity across the filter curves, and/or the onset of complete GP-troughs at
z>5.7?

This is Figure 2, not 3.
And I think this is where some of the missing objects come in??
We’ll need to update: Tables 1, 2 and obviously 3 once we have those extra objects.

p. 4, Fig. 3: Can you plot the small dots somewhat bigger for better visibility
without loss of information?

This is the coverage map figure
       => To Do by NJGC.

THROUGHOUT: Several occurrences of ??, i.e., where table or Appendix references
are not properly done in latex.

NPR to double check…

p. 5: [My only (minor) scientific comment]: Re: ``although possible, we consider
it highly unlikely there would be more than one MIR source contributing to the
flux of a single UV/optically bright rest-frame quasar. Investigating this in
detail is very interesting but left to a future study. ''

I think it would be better to address now the number of expected contaminating
8--25 micron sources in the apertures used, as it can be done straightforwardly.
E.g., Driver et al. (2016, ApJ, 827, 107; astro-ph/1605.01523) --- the figures
you need for this estimate are actually best given as Fig. A4, A5 on the astro-ph
version. The redshift distribution of the sampled 8--25 micron foreground objects
is narrow enough and their K-corrections advantageous enough, that the objects
counts sharply turnover for AB(W3, W4) > 20--18 mag (left panels), and as a
consequence the normalized differential counts (right panels) --- especially in
W4 --- have most of their EBL energy emerging at AB~18--20 mag. As a consequence,
the relative number or probability of contaminating objects at AB>20 mag inside
your apertures should be rather small. A point worth mentioning, as it helps your
case.

NPR to add this to the text.

p. 6 Table 3: Should mention that the number of significant digits listed here
for the YJHK W1--W4 mags is not always warranted, but needed to verify the small
VIRCAM-WFCAM offsets in Table 5, This table is a very useful check of your global
photometric zeropoints.

NPR to change.

p. 7, very end of 3.1.1: Quote the (effective) VSA and SDSS quasar IR observing
epochs here, in case these clearly discrepant magnitudes are due to variability.

This is for SDSS J0349+0034 and J2220-0101; NJGC to check.

p. 8 Fig 5: Make the z=0--3 quasar cloud a little lighter grey, so that the
z=5--6.5 quasar cloud is better visible.

NPR to update.

p. 9 top left: = -0.458 : All minus signs in latex should be coded up as --
(double hyphen), also in the tables, to appear as minus sign.

NPR to update.

p. 10 Fig 8: Enlarge all axes label for better readability.

NPR to update.

THROUGHOUT: signal-to-noise should always be listed as a ratio: signal-to-noise
ratio or SNR.
NPR to update.