### 2020-01-13 tex/MNRAS_submission/rv1/referees_reply.txt
Fixed numbers for VegaToAB, added to referee reply

	We thank the Referee for reading our manuscript carefully and
	for their review and comments.  We address each and every one
	of these in turn below (indented text).

	We have also made further, key improvements to our dataset and
	pipeline.

	We note that after the submission of the original manuscript
	it was brought to our attention (K. Perger; priv. comm) that
	we were missing 25 confirmed z>5 quasars.  As such, our
	database has increased by 25 objects for a total of 488 VHzQs.

	As such all six Tables have been updated accordingly, as have
	Figures 2 and 3.

	We also fully re-ran the WFCAM Science Archive (WSA) and VISTA
	Science Archive (VSA) from scratch with the set of updated
	objects. This uncovered a small matching bug for three objects
	in the WSA (no issues were found in the VSA rerun). One of
	these objects, however, is SDSS J2220-0101, which we thought
	had large photometric changes but now acknowledge this was due
	to an incorrect apertureID assignment, when the database was updated 
	with new data. The updated photometry for SDSS J2220-0101 has the  
	VSA J-band magnitude 20.32+/-0.07 mag AB and a WSA magnitude 20.30+/-0.15 mag AB.

	Consequently, we have changed the appropriate text (in the Abstract,
	Section 3.1.1 Comparing WFCAM and VISTA and Conclusions).

	We have also added an Appendix to note the update to the overall catalog.

p. 1: You may want to end the first sentence of the right column of the Intro
with a statement that at z>5, relativistic time dilation also stretches the
photometric variability of the accretion disk as observed.

	    Now added.

p. 2, Fig. 1: Is it useful here to plot both filter sets if they are sufficiently
different? Please provide larger tick-marks for better readability in any case.

	   We have enbiggened the  tick-marks.

	   The ZYJHK filters for UKIRT and VISTA *are* different, but
	   plotting both doesn't result in a clear figure.

	   ....
		I think adding an explicit link to:
	          d80b2t/VHzQ/SEDs/filters_vs_QSOstars_UKIRT.png
	  and
	         d80b2t/VHzQ/SEDs/filters_vs_QSOstars_VISTA.png
	 is the way forward here.

p. 3, Fig. 2, caption or text: You should briefly comment on why the quasar
distribution has this gap in N(z) at z~5.5?. Is this because of the relativity
sensitivity across the filter curves, and/or the onset of complete GP-troughs at
z>5.7?

	[And I think this is where some of the missing objects come in??
	We’ll need to update: Tables 1, 2 and obviously 3 once we have those extra objects.]

p. 4, Fig. 3: Can you plot the small dots somewhat bigger for better visibility
without loss of information?

	We have updated the plot to make the dots bigger and the revised catalogue.

THROUGHOUT: Several occurrences of ??, i.e., where table or Appendix references
are not properly done in latex.

    	Checked and updated.

p. 5: [My only (minor) scientific comment]: Re: ``although possible, we consider
it highly unlikely there would be more than one MIR source contributing to the
flux of a single UV/optically bright rest-frame quasar. Investigating this in
detail is very interesting but left to a future study. ''

I think it would be better to address now the number of expected contaminating
8--25 micron sources in the apertures used, as it can be done straightforwardly.
E.g., Driver et al. (2016, ApJ, 827, 107; astro-ph/1605.01523) --- the figures
you need for this estimate are actually best given as Fig. A4, A5 on the astro-ph
version. The redshift distribution of the sampled 8--25 micron foreground objects
is narrow enough and their K-corrections advantageous enough, that the objects
counts sharply turnover for AB(W3, W4) > 20--18 mag (left panels), and as a
consequence the normalized differential counts (right panels) --- especially in
W4 --- have most of their EBL energy emerging at AB~18--20 mag. As a consequence,
the relative number or probability of contaminating objects at AB>20 mag inside
your apertures should be rather small. A point worth mentioning, as it helps your
case.

		We have added approriated words in the text to address this good point.

p. 6 Table 3: Should mention that the number of significant digits listed here
for the YJHK W1--W4 mags is not always warranted, but needed to verify the small
VIRCAM-WFCAM offsets in Table 5, This table is a very useful check of your global
photometric zeropoints.

p. 7, very end of 3.1.1: Quote the (effective) VSA and SDSS quasar IR observing
epochs here, in case these clearly discrepant magnitudes are due to variability.

SDSS J0349+0034 was observed on MJD=54062.4 by WFCAM and MJD=55540.6 (average of 
3 epochs from 55537 to 55542) by VISTA. We have updated the text with these values.
We have also updated the values to AB magnitudes, as we found that we had inadvertently 
given the Vega magnitude values.

p. 8 Fig 5: Make the z=0--3 quasar cloud a little lighter grey, so that the
z=5--6.5 quasar cloud is better visible.

p. 9 top left: = -0.458 : All minus signs in latex should be coded up as --
(double hyphen), also in the tables, to appear as minus sign.

p. 10 Fig 8: Enlarge all axes label for better readability.

THROUGHOUT: signal-to-noise should always be listed as a ratio: signal-to-noise
ratio or SNR.